Maximising the guaranteed feasible set for uncertain MPC schemes with probabilistic constraints

### 2016-11-05 draft.tex
edits 2.1

\providecommand{\conv}{\text{conv}}
\providecommand{\vol}{\text{vol}}
\providecommand{\spann}{\text{span}}
\newcommand{\Obs}{\mathcal{O}}
\providecommand{\B}{\mathcal B}
\providecommand{\C}{\mathcal C}
\providecommand{\E}{\mathcal E}
\providecommand{\W}{\mathcal W}
\providecommand{\V}{\mathcal V}
\providecommand{\X}{\mathcal X}
\providecommand{\Y}{\mathcal Y}
\providecommand{\Z}{\mathcal Z}
\providecommand{\U}{\mathcal U}
\providecommand{\R}{\mathcal R}
\providecommand{\T}{\mathcal T}
\providecommand{\D}{\mathscr D}
\providecommand{\PP}{\mathbb P}
\providecommand{\RR}{\mathbb R}
\providecommand{\bfa}[1]{\mathbf{#1}}

@@

%\allowdisplaybreaks[1]

@@

\begin{abstract}                          % Abstract of not more than 250 words.
This paper proposes a method of approximating positively invariant sets and $n$-step controllable sets of uncertain linear systems that are subject to chance constraints. The computed sets are robustly invariant and are guaranteed to satisfy the probabilistic constraints; in contrast existing methods based on random sampling are only able to satisfy such constraints with a certain level of confidence. The proposed approach uses explicitly parametrised auxiliary disturbance sets, which are optimised so as to maximise the relevant positively invariant or $n$-step controllable set. The results are illustrated by numerical examples.
\end{abstract}

@@

The following notation is employed: $A\oplus B = \{x:x=a+b,\, a\in A,\, b\in B\}$ denotes the Minkowski addition, $A\ominus B = \{x: x+b \in A \, \forall b\in B\}$ the Pontryagin difference, for a real set $C\subset\RR^d$ we denote the image of $C$ under the linear map $A:\RR^d\rightarrow\RR^d$ by $AC = \{y\in\RR^d:Ax=y,x\in C\}$ and the pre-image $A^{-1}C = \{y\in\RR^d:Ay\in C\}$.
%
We denote the probability of an event $E$ by $\PP\{E\}$, throughout this paper we assume the probability density to be uniform.
%
The set $\B_P(r) = \{x\in\RR^d:x^TPx\leq r^2\}$ denotes the $P$-ball of radius $r$ centred at the origin, for a matrix $\Gamma\in\RR^{m\times n}$ the notation $\Gamma x\leq \bfa{1}$ is to be understood element-wise, where $\bfa{1}$ is a vector of ones in $\RR^m$.

@@

The discrete time system we consider is given by
%
\begin{equation}
	x^+ = Ax + Bu + w
\end{equation}
%
with state $x$ and control input $u$ subject to the constraints 
\begin{align*}
&x \in\X = \bigl\{x\in\RR^d : E_ix\leq 1, \ i\leq M_\X\bigr\}, \\
&u\in\U = \bigl\{u\in\RR^{q_\U} : F_i u\leq 1, \ i\leq M_\U\bigr\},
\end{align*} 
and with disturbance input $w$ subject to the constraints
\[
w\in\W = \bigl\{w\in\RR^{q_\W}:G_i w\leq 1,\ i\leq M_\W \bigr\}.
\]
%
We furthermore consider the auxiliary output variable 
%
\begin{equation}
	y = Cx + Du + v
\end{equation}
%
which depends on a random variable $v$:
\[
v\in\V = \bigl\{v\in\RR^{q_\Y}:\Gamma_i v\leq 1, \ i\in M_\V\bigr\},
\]
for which the probability distribution is assumed to be uniform.
%
The output is probabilistically constrained to satisfy
%
\begin{equation}\label{eq:probabilistic:constraint}
	\PP\bigl\{ v \in \mathcal{V} : y = Cx + Du + v \in \mathcal{Y} \bigr\}\geq p.
\end{equation}
%
with
\[
\Y=\bigl\{y\in\RR^{q_\Y} : H_i y\leq h_i, \ i\leq M_\Y\bigr\}.
\]

@@

\vspace{0.5\baselineskip}\begin{prop}
Let $v\in\V$ be a uniformly distributed random variable, let $\V^\prime\subseteq\V$ such that $\vol(\V^\prime)\geq p$, then every $q\in\Z\ominus\V^\prime$ satisfies $\PP\{q+v\in\Z\}\geq p$.
\end{prop}
%
\bpf
%
Using the fact that $\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} = \PP\{\V^\prime\} = \vol(\V^\prime)$ for all $q\in\Z\ominus\V^\prime$ we have
%
\begin{equation}\begin{split}
&\PP\{q+v\in\Z\} = \PP\{\{q\}\oplus\V\subseteq\Z\} 
 \\
&=\PP\{\{q\}\oplus(\V^\prime\cup(\V\setminus\V^\prime))\subseteq\Z\} \\
&=\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} + \PP\{\{q\}\oplus(\V\setminus\V^\prime)\subseteq\Z\} \\
&\geq\PP\{\{q\}\oplus\V^\prime\subseteq\Z\}\\ &=\vol(\V^\prime)\geq p.
\end{split}\end{equation}
\mbox{}\vspace{-0.5\baselineskip}
\epf
%

Notice that for polytopic $\Z$ and $\V^\prime$ the set $\Z\ominus\V^\prime$ can be computed explicitly and is itself polytopic.

In the sequel of this paper we discuss how~$\V^\prime$ can be chosen in order to obtain the largest guaranteed n-step controllable set for a given target set, guaranteed positive invariant set.

\subsection{Approximating the guaranteed invariant set}\label{ssec:approximating:MRPI}
%
First we discuss how we can extend existing methods to calculate the maximal robust positive invariant (MRPI) set (see e.g.~\cite{blanchini:2007}) in such a way that all states produce only feasible successor states for a given feedback controller $u = Kx$.
%
I.e. an invariant set for which all possible trajectories satisfy state constraints and probabilistic constraint at all time.
%
\begin{multline}\label{eq:alternative:MRPI:set}
	X^\infty = \Bigl\{x\in\X: \\
\begin{aligned} 
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i} \\
&x_k\in\X\\
&(C+DK)x_k+v\in\Y
\end{aligned}\\
\forall w_j\in\W,\ \forall v\in \V^\prime,\ \forall k\geq 0 \Bigr\}
\end{multline}
% \begin{multline}\label{eq:alternative:MRPI:set}
% 	X^\infty = \Bigl\{x\in\X: \\
% \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,\ \forall v\in \V^\prime,\ \forall k\geq 0 \Bigr\}
% \end{aligned} 
% \end{multline}
%
It is obvious that $X^\infty$ contained in the invariant set of the original problem~$\X^\infty$:
%
\begin{multline*}
\X^\infty = \Bigl\{x\in\X: \\
\begin{aligned}
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
&x_k\in\X\\
&\PP\{(C+DK)x_k+v\in\Y\}\geq p 
\end{aligned}\\
\forall w_j\in\W,\, \forall k\geq 0 \Bigr\}
\end{multline*}
% \begin{equation}
% 	\X^\infty = \left\{x\in\X: \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}
% \end{equation}
%
The inclusion $X^\infty\subseteq\X^\infty$ holds for any $\V^\prime\subseteq\V$ with the appropriate probability measure, therefore maximising the volume of $X^\infty$ improves the approximation of $\X^\infty$.

In order to determine an algorithm to compute $X^\infty$ we rewrite the conditions met in~\eqref{eq:alternative:MRPI:set} we introduce the set of all possible successor states~$\D_k(X)$ of states in $X$ at time $k$:
%
\begin{align*}
	\D_k(X) &= (A+BK)^{k}X\oplus\W\oplus\dots\oplus(A+BK)^{k-1}\W\\
	&= (A+BK)^k X\oplus \bigoplus_{i=0}^{k-1}(A+BK)^i\W
\end{align*}
%
Using $\R = (C+DK)^{-1}(\Y\ominus\V^\prime)$ we can abbreviate condition~\eqref{eq:alternative:MRPI:set} to
%
\begin{equation}\label{eq:containment:condition:alternative:MRPI}
\D_k(X^\infty)\subseteq\X\cap\R
\end{equation}
%
for all $k\geq0$.
%
With this identity we can determine a direct way of determining $X^\infty$:
%
\begin{gather*}
X^\infty = %\bigcap_{k\geq0}(A+BK)^{-k}\Bigl((\X\cap\R)\ominus\bigoplus_{i=0}^{k-1}(A+BK)^i\W\Bigr) \\
\bigcap_{k\geq0} \E_k, \\
\E_k = (A+BK)^{-k}\Bigl((\X\cap\R)\ominus\bigoplus_{i=0}^{k-1}(A+BK)^i\W\Bigr) .
\end{gather*}
% \begin{equation}\label{eq:direct:formulation:MRPI:with:intersection}
% 	X^\infty = \bigcap_{k\geq0}\underbrace{(A+BK)^{-k}\left((\X\cap\R)\ominus\bigoplus_{i=0}^{k-1}(A+BK)^i\W\right)}_{\E_k}.
% \end{equation}
%
Due to its definition~$\E_k$ satisfies $\D_k(\E_k)\subseteq\X\cap\R$ for each $k\geq0$, furthermore using available methods it can be shown that $\E_k$ expands exponentially and therefore $X^\infty$ is finitely determined.
%
In order to computationally determine $X^\infty$ we define the recursive sequence
%
\begin{equation}\label{eq:MRPI:sequence}
	\begin{split}
	X_0 &= \X\\
	Z_k &= X_k\cap\R\\
	X_{k+1} &= Z_k\cap(A+BK)^{-1}(Z_k\ominus\W)
	\end{split}
\end{equation}
%
Notice that $X_k = \bigcap_{l\leq k}\E_l$ and therefore the recursion terminates, analogue to the conventional MRPI calculation, once $X_k\subseteq X_{k+1}$.

\vspace{0.5\baselineskip}
\begin{thm}
Let $\Gamma$ be a map such that $\Gamma x\leq\bfa{1}$ and $-\Gamma x\leq\bfa{1}$ for all $x\in\X$ and let $\bigl((A+BK),\Gamma\bigr)$ be observable, then iteration~\eqref{eq:MRPI:sequence} terminates after a finite number of iterations~$N$, such that $X_{N+1}=X_N$ and $X_N=X^\infty$.
\end{thm}
%
\bpf
%
The observability of $\bigl((A+BK),\Gamma\bigr)$ implies that $X_d$ is compact, since 
\begin{align*}
X_{d-1} 
&\subseteq\bigcap_{k\leq d-1}(A+BK)^{-k}\X\\
&\subseteq \Bigl\{x:\pm\Gamma x\leq\bfa{1}\wedge\dots\wedge\pm\Gamma(A+BK)^{d-1} x\leq\bfa{1} \Bigr\}\\
& = \{x:\pm\Gamma \Obs x\leq \bfa{1}\}
\end{align*}
where $\Obs$ is the (full rank) observability matrix.

Let~$P$ be such that 
\[
(A+BK)^TP(A+BK) \leq \rho^2P
\]
holds, where $\rho$ denotes the spectral radius of $A+BK$.
%
To see that $\E_k$ grows exponentially let $\W\subseteq\B_P(r)$ then
%
\begin{multline*}
  \bigoplus_{i=0}^{k-1}(A+BK)^i\W \subseteq \bigoplus_{i=0}^{k-1}(A+BK)^i\B_P(r)\\
  \subseteq\bigoplus_{i=0}^{k-1}\B_P(\rho^i r) = \B_P(\sum_{i=0}^{k-1}\rho^i r)\subseteq\B_P(\frac{r}{1-\rho})
\end{multline*}
%
i.e. $\E_k$ consists of an expanding minuend and a bounded subtrahend.
%
Therefore $\E_k$ expands exponentially, since $X_{d-1}$ is bounded there exists an integer~$N$ such that $X_{N}\cap\E_{N+1}=X_N$ hence $X_N=X^\infty$.
%
\epf

The optimisation program to approximate the positive invariant set $\X^\infty$ can therefore be formulated as
%
\begin{subequations}\label{seq:optimisation:MRPI:abstract}
	\begin{equation}
		\max_{\V^\prime} \vol(X^\infty)
	\end{equation}
%
	subject to
%
\begin{equation}
	\V^\prime\subseteq\V
\end{equation}
%
\begin{equation}
	\vol(\V^\prime)\geq p\cdot\vol(\V)
\end{equation}
\end{subequations}
%
In section~\ref{sec:volume:of:hypercube} we will discuss how an optimisation can be solved using parallelotopes for~$\V^\prime$.

\subsection{Approximating the n-step controllable set}\label{ssec:approximating:n:step}
%
We now turn our attention to the n-step controllable set~$\C_n(\T)$ for a given target set~$\T$.
%
That is the set of all states such that a sequence of admissible inputs exists that takes any state $x\in\C_n$ to the target set $x_n\in\T$ for any admissible disturbance sequence $w\in\W$ while guaranteeing state- and probabilistic constraint satisfaction.
%
This set is given by
%
\begin{multline*}
\C_n(\T) = \Bigl\{x\in\X: \\
\begin{aligned}
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
&x_N\in\T\\
&x_k\in\X\\
&\PP\{(C+DK)x_k+v\in\Y\}\geq p
\end{aligned}\\
\forall w_j\in\W,\, \forall k\geq 0 \Bigr\}	
\end{multline*}
% \begin{equation}
% 	\C_n(\T) = \left\{x\in\X:\begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_N\in\T\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}	
% \end{equation}
%
Again, it is trivial to see that the set
%
\begin{multline*}
C_n(\T) = \Bigl\{ x\in\X: \\
\begin{aligned}
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
&x_N\in\T\\
&x_k\in\X\\
&(C+DK)x_k+v\in\Y
\end{aligned}\\
\forall w_j\in\W,\, \forall v\in \V^\prime,\, \forall k\in\{0,\dots,N\} \Bigr\}	
\end{multline*}
% \begin{equation}
% 	C_n(\T) = \left\{x\in\X:\begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_N\in\T\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,v\in \V^\prime,k\in\{0,\dots,N\}\end{aligned} \right\}	
% \end{equation}
%
is always contained in $\C_n$ for any admissible choice of $\V^\prime\subseteq\V$.
%
As in the case of the positive invariant set, fixing $\V^\prime$ to be polytopic leads to a computationally tractable algorithm yielding~$C_n$ to be a polytopic approximation of~$\C_n$, and optimising over~$\V^\prime$ to maximise the volume of~$C_n$ indirectly improves the approximation of $\C_n$.
%
It is known that $C_n(\T) = C^n_1(\T) = C_1(C_1(\dots C_1(\T)\dots))$ (see e.g.~\cite{blanchini:2007}), which allows the use of simple computational geometry algorithms.

>>>>

\providecommand{\conv}{\text{conv}}
\providecommand{\vol}{\text{vol}}
\providecommand{\spann}{\text{span}}
\newcommand{\Obs}{\mathcal{O}}
\providecommand{\B}{\mathcal B}
\providecommand{\C}{\mathcal C}
\providecommand{\E}{\mathcal E}
\providecommand{\F}{\mathscr F}
\providecommand{\W}{\mathcal W}
\providecommand{\V}{\mathcal V}
\providecommand{\X}{\mathcal X}
\providecommand{\Y}{\mathcal Y}
\providecommand{\Z}{\mathcal Z}
\providecommand{\U}{\mathcal U}
\providecommand{\R}{\mathcal R}
\providecommand{\T}{\mathcal T}
\providecommand{\D}{\mathscr D}
\providecommand{\PP}{\mathbb P}
\providecommand{\RR}{\mathbb R}
\providecommand{\bfa}[1]{\mathbf{#1}}

@@

\allowdisplaybreaks[4]

@@

\begin{abstract}                          % Abstract of not more than 250 words.
This paper proposes a method of approximating positively invariant sets and $n$-step controllable sets of uncertain linear systems that are subject to chance constraints. The computed sets are robustly invariant and are guaranteed to satisfy the probabilistic constraints. In contrast, existing methods based on random sampling are only able to satisfy such constraints with a certain level of confidence. The proposed approach uses explicitly parametrised auxiliary disturbance sets, which are optimised so as to maximise the relevant positively invariant or $n$-step controllable set. The results are illustrated by numerical examples.
\end{abstract}

@@

\textit{Notation.} Given sets $A,B\subset\RR^d$, $A\oplus B = \{x:x=a+b,\, a\in A,\, b\in B\}$ denotes the Minkowski sum and $A\ominus B = \{x: x+b \in A \ \forall b\in B\}$ the Pontryagin difference. We denote the image and pre-image of $C\subset\RR^d$ under the linear map $A:\RR^d\rightarrow\RR^d$ respectively as $AC = \{y\in\RR^d:Ax=y,\, x\in C\}$ and $A^{-1}C = \{y\in\RR^d:Ay\in C\}$. The $d$-dimensional volume of $C\subset\RR^d$ is denoted by $\vol(C)$. 
%
We define $v$ as a random parameter on a probability space $(\V,\F,\PP)$ and denote the probability of an event $E(v)$ as $\PP\{v\in\V : E(v)\}$.
%; throughout this paper the probability density of $v$ is assumed to be uniform.
%
The set $\B_P(r) = \{x\in\RR^d:x^TPx\leq r^2\}$ denotes the $P$-ball of radius $r$ centred at the origin. For a matrix $\Gamma\in\RR^{m\times n}$ the notation $\Gamma x\leq \bfa{1}$ is to be understood element-wise, where $\bfa{1}$ is a vector of ones in $\RR^m$.

@@

The discrete time system we consider is given by
%
\begin{equation}\label{eq:system}
	x^+ = Ax + Bu + w ,
\end{equation}
%
where the state $x$ and control input $u$ are subject to constraints defined in terms of compact, convex polytopic sets $\X$, $\U$:
\begin{equation}\label{eq:state_input:constraints}
\begin{aligned}
&x \in\X = \bigl\{x\in\RR^d : E_ix\leq 1, \, i\leq M_\X\bigr\}, \\
&u\in\U = \bigl\{u\in\RR^{q_\U} : F_i u\leq 1, \, i\leq M_\U\bigr\},
\end{aligned}
\end{equation} 
and the disturbance input $w$ is subject to the constraint
\begin{equation}\label{eq:disturbance}
w\in\W = \bigl\{w\in\RR^{q_\W}:G_i w\leq 1,\, i\leq M_\W \bigr\}.
\end{equation}
%
We furthermore consider the output variable $y$,
%
\[%\begin{equation}\label{eq:output}
	y = Cx + Du + v
\]%\end{equation}
%
which depends on the random variable $v$,
\[
v\in\V = \bigl\{v\in\RR^{q_\Y}:\Gamma_i v\leq 1, \ i\in M_\V\bigr\}. 
\]
Throughout this paper $v$ is assumed to be uniformly distributed on $\V$.
%Throughout this paper the probability density function of $v$ is assumed to be uniform.
%
The output variable $y$ is subject to the chance (or probabilistic) constraint
%
\begin{equation}\label{eq:probabilistic:constraint}
	\PP\bigl\{ v \in \mathcal{V} : y = Cx + Du + v \in \mathcal{Y} \bigr\}\geq p
\end{equation}
%
for given $p\in(0,1]$, with
\[
\Y=\bigl\{y\in\RR^{q_\Y} : H_i y\leq h_i, \ i\leq M_\Y\bigr\}.
\]

@@

\vspace{0.5\baselineskip}
\begin{prop}\label{prop1}
Let $v\in\V$ be a uniformly distributed random variable and let $\V^\prime\subseteq\V$ be such that $\vol(\V^\prime)\geq p$, then every $q\in\Z\ominus\V^\prime$ satisfies $\PP\{v\in\V : q+v\in\Z\}\geq p$.
\end{prop}
%

\bpf
%
For any given $q\in\Z\ominus\V^\prime$, the definition of the Pontryagin difference implies that
\[
q+v\in\Z \quad \text{whenever} \quad  v\in\V^\prime.
\]
Therefore, for every $q\in\Z\ominus\V^\prime$, the probability of the event that $q+v$ lies in $\Z$ can be no less than the probability of the event that $v$ lies in $\V^\prime$. It follows that 
\[
\PP\{ v\in\V : q + v\in \Z\} \geq \PP\{\V^\prime\} = \vol(\V^\prime)
\]
and hence $\PP\{ v\in\V : q + v\in \Z\} \geq p$ if $\V^\prime$ is such that $\vol(\V^\prime) \geq p$.
% Using the fact that $\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} = \PP\{\V^\prime\} = \vol(\V^\prime)$ for all $q\in\Z\ominus\V^\prime$ we have
% %
% \begin{align*}
% &\PP\{v\in\V : q+v\in\Z\} 
% =\PP\{v\in \V^\prime\cup(\V\setminus\V^\prime) : q + v \in \Z, \ ()\subseteq\Z\} \\
% &=\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} + \PP\{\{q\}\oplus(\V\setminus\V^\prime)\subseteq\Z\} \\
% &\geq\PP\{\{q\}\oplus\V^\prime\subseteq\Z\}\\ &=\vol(\V^\prime)\geq p. 
% \end{align*}
%\mbox{}\vspace{-0.5\baselineskip}
\epf
%

Note that for polytopic $\Z$ and $\V^\prime$ the set $\Z\ominus\V^\prime$ can be computed explicitly and is itself polytopic.

In the remainder of this paper we discuss how~$\V^\prime$ can be chosen in order to obtain the largest guaranteed $n$-step controllable set for a given target set and the largest guaranteed positively invariant set for the system (\ref{eq:system}) subject to the robust constraints~(\ref{eq:state_input:constraints}) and the chance constraint (\ref{eq:probabilistic:constraint}).

\subsection{Approximating the guaranteed invariant set}\label{ssec:approximating:MRPI}
%
First we discuss how to extend existing methods to calculate the maximal robust positively invariant (MRPI) set (see e.g.~\cite{blanchini:2007}) in such a way that all states in the MRPI set produce only feasible successor states for a given feedback controller $u = Kx$.
%
Thus we seek to determine a set of initial conditions for which all possible trajectories of (\ref{eq:system}) subject to (\ref{eq:disturbance}) satisfy the robust constraints~(\ref{eq:state_input:constraints}) and probabilistic constraint (\ref{eq:probabilistic:constraint}). Denoting the largest such set as $\X^\infty$ we have
\begin{multline*}
\X^\infty = \Bigl\{x\in\X: \\
\begin{aligned}
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i} w_{k-i}\\
&x_k\in\X\\
&Kx_k\in\U\\
&\PP\{v\in\V : (C+DK)x_k+v\in\Y\}\geq p 
\end{aligned}\\
\forall w_k\in\W,\, \forall k\geq 0 \Bigr\} .
\end{multline*}
%
% \begin{multline}\label{eq:alternative:MRPI:set}
% 	X^\infty = \Bigl\{x\in\X: \\
% \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,\ \forall v\in \V^\prime,\ \forall k\geq 0 \Bigr\}
% \end{aligned} 
% \end{multline}
%
% It is obvious that $X^\infty$ contained in the invariant set of the original problem~$\X^\infty$:
For a given set $\V^\prime$, let $X^\infty$ denote the set
%
\begin{multline}\label{eq:alternative:MRPI:set}
	X^\infty = \Bigl\{x\in\X: \\
\begin{aligned} 
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i} w_{k-i} \\
&x_k\in\X \\ 
&Kx_k\in\U \\
&(C+DK)x_k+v_k\in\Y
\end{aligned}\\
\forall w_k\in\W,\ \forall v_k\in \V^\prime,\ \forall k\geq 0 \Bigr\} .
\end{multline}
% \begin{equation}
% 	\X^\infty = \left\{x\in\X: \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}
% \end{equation}
%
Then from Proposition~\ref{prop1}, the inclusion $X^\infty\subseteq\X^\infty$ necessarily holds for any $\V^\prime\subseteq\V$ with probability measure $\PP\{\V^\prime\} \geq p$. Furthermore, by maximising the volume of $X^\infty$, we can improve the approximation of $\X^\infty$.

In order to determine an algorithm to compute $X^\infty$ we rewrite the conditions in~\eqref{eq:alternative:MRPI:set} by introducing the set of all possible successor states~$\D_k(X)$ of states in $X$ after $k$ time steps:
%
\begin{align*}
	\D_k(X) &= (A+BK)^{k}X\oplus\W\oplus\dots\oplus(A+BK)^{k-1}\W\\
	&= (A+BK)^k X\oplus \bigoplus_{i=0}^{k-1}(A+BK)^i\W
\end{align*}
%
Defining $\R = (C+DK)^{-1}(\Y\ominus\V^\prime)$ we can abbreviate the conditions defining $X^\infty$ in~\eqref{eq:alternative:MRPI:set} to
%
\begin{equation}\label{eq:containment:condition:alternative:MRPI}
\D_k(X^\infty)\subseteq\X\cap\R
\end{equation}
%
for all $k\geq0$.
%
Using these conditions we can express $X^\infty$ directly as
%
\begin{gather}
X^\infty = %\bigcap_{k\geq0}(A+BK)^{-k}\Bigl((\X\cap\R)\ominus\bigoplus_{i=0}^{k-1}(A+BK)^i\W\Bigr) \\
\bigcap_{k\geq0} \E_k, \\
\E_k = (A+BK)^{-k}\biggl[(\X\cap\R)\ominus\bigoplus_{i=0}^{k-1}(A+BK)^i\W\biggr] . \label{eq:Ek:subtract}
\end{gather}
% \begin{equation}\label{eq:direct:formulation:MRPI:with:intersection}
% 	X^\infty = \bigcap_{k\geq0}\underbrace{(A+BK)^{-k}\left((\X\cap\R)\ominus\bigoplus_{i=0}^{k-1}(A+BK)^i\W\right)}_{\E_k}.
% \end{equation}
%
Due to its definition~$\E_k$ satisfies $\D_k(\E_k)\subseteq\X\cap\R$ for each $k\geq0$, furthermore it can be shown that $\E_k$ expands exponentially and therefore $X^\infty$ is finitely determined  (see Theorem~\ref{thm2} below).
%
In order to compute $X^\infty$ we define the sequence $X_0,X_1,\ldots$ via the iteration:
%
\begin{equation}\label{eq:MRPI:sequence}
\begin{aligned}
  & X_0 = \X \\
  & \begin{aligned}
      & Z_k = X_k\cap\R\\
      & X_{k+1} = Z_k\cap(A+BK)^{-1}(Z_k\ominus\W)
    \end{aligned} 
  \ \Biggr\} \ \ k = 1,2,\ldots
\end{aligned}
\end{equation}
%
%with $X_0= \X$.
Clearly $X_k = \bigcap_{l=0,\ldots,k}\E_l$ and, analogously to the conventional MRPI calculation~\citep{Kolmanovsky:1998}, the iteration therefore gives $X_k = X^\infty$ once $X_k\subseteq X_{k+1}$.

\vspace{0.5\baselineskip}
\begin{thm}\label{thm2}
Let $\Gamma$ be a matrix such that $\Gamma x\leq\bfa{1}$ and $-\Gamma x\leq\bfa{1}$ for all $x\in\X$ and let $\bigl((A+BK),\Gamma\bigr)$ be observable, then the iteration~\eqref{eq:MRPI:sequence} %terminates after a finite number, $N$, of iterations, such that 
yields $X_{k+1}=X_k$ for all $k\geq N$ and hence $X_k=X^\infty$ for some finite integer $N$.
\end{thm}
%
\bpf
%
The observability of the pair $\bigl((A+BK),\Gamma\bigr)$ implies that $X_{d-1}$ is compact, since 
\begin{align*}
X_{d-1} 
&\subseteq\bigcap_{k\leq d-1}(A+BK)^{-k}\X\\
&\subseteq \Bigl\{x:\pm\Gamma x\leq\bfa{1}\wedge\dots\wedge\pm\Gamma(A+BK)^{d-1} x\leq\bfa{1} \Bigr\}\\
& = \bigl\{x:\pm\Gamma \Obs x\leq \bfa{1}\bigr\}
\end{align*}
where $\Obs$ is the (full rank) observability matrix.

Let~$P$ be such that 
\[
(A+BK)^TP(A+BK) \leq \rho^2P
\]
holds, where $\rho$ denotes the spectral radius of $A+BK$.
%
To see that $\E_k$ grows exponentially let $\W\subseteq\B_P(r)$ then
%
\begin{multline*}
  \bigoplus_{i=0}^{k-1}(A+BK)^i\W \subseteq \bigoplus_{i=0}^{k-1}(A+BK)^i\B_P(r)\\
  \subseteq\bigoplus_{i=0}^{k-1}\B_P(\rho^i r) = \B_P\biggl(\sum_{i=0}^{k-1}\rho^i r\biggr)\subseteq\B_P\Bigl(\frac{r}{1-\rho}\Bigr) .
\end{multline*}
%
It follows that the subtrahend in square brackets on the RHS of (\ref{eq:Ek:subtract}) is bounded for all $k$ by $\B_P\bigl(r / (1-\rho)\bigr)$.
%It follows that the RHS of (\ref{eq:Ek:subtract}) consists of an expanding minuend and a bounded subtrahend.
%
Therefore 
%$\E_k$ expands exponentially. Hence 
the iteration (\ref{eq:MRPI:sequence}) must either yield $X_N = \emptyset$ for finite $N$
(in which case $X^\infty=\emptyset$) or,
since $X_{d-1}$ is bounded and $\E_k$ expands exponentially, there must exist a finite $N$ such that $X_{N}\cap\E_{N+1}=X_N$ and hence $X_N=X^\infty$.
%
\epf

An optimisation program to approximate the positively invariant set $\X^\infty$ can therefore be formulated as
%
\begin{subequations}\label{seq:optimisation:MRPI:abstract}
	\begin{equation}
		\max_{\V^\prime} \vol(X^\infty)
	\end{equation}
%
	subject to
%
\begin{equation}
	\V^\prime\subseteq\V
\end{equation}
%
\begin{equation}
	\vol(\V^\prime)\geq p\cdot\vol(\V)
\end{equation}
\end{subequations}
%
In section~\ref{sec:volume:of:hypercube} we discuss how this optimisation can be solved using parallelotopes for~$\V^\prime$.
